# Counts and Tables {#tables}


:::note
You should use this method if the data is:

- Categorical
:::

In this chapter you will learn how to do some simple data explorations for categorical variables using tables (also called  tables) and simple counts. 


## Simple counts 

Get an overview of the variable `maritl` and sort the values. We use the function gt() to print nice tables:

```{r count-one}

wage_df %>% 
  count(maritl,
  sort = TRUE) %>% 
  gt()

```

Get an overview of the combined variables `maritl` and `education`:

```{r count-two}

wage_df %>% 
  count(maritl, education) %>% 
   gt()

```

Obtain the sum of a quantitative variable (`wage`) for the different levels of a categorical variable (`maritl`):

```{r count-numeric}

wage_df %>%  
  count(maritl,
        wt = wage,
        name = "Wage_sum") %>% 
  gt()

```

## Total counts

Total counts are an useful way to represent the observations that fall into each combination of the levels of categorical variables. We create a contingency table of the two categorical variables `jobclass` and `race` and call the result `tab`:

```{r tab}

tab <- table(wage_df$jobclass, wage_df$race)

```

## Joint proportions

We can also view the percentage of each cell in relation to the total amount of all observations (here n = 3000). Therefore, you have to simply divide the numbers from our total counts with 3.000.

The following code generates tables of *joint* proportions:

```{r prop-tab}

# joint proportions
prop.table(tab)

```

For example, around 44% of all people in the dataset are white industrial workers.


## Conditional proportions: columns

You also may want to know the probability that workers have a certain jobclass, given that they have a particular ethnical background. This is a so called conditional probability. Conditional probability represents the chance that one event will occur given that a second event has already occurred. 

The following code generates tables of *conditional* proportions:

```{r prop-tab-col}

# conditional on columns
prop.table(tab, 2)  

```

We performed a columnwise evaluation and are now able to answer the following question: 

* Approximately what proportion of all white workers are industrial workers? 
* The answer is: around 53%. 

## Conditional proportions: rows

Now we want to obtain the probability that workers have a certain race, given their jobclass. 


```{r prop-tab-row}

# conditional on rows
prop.table(tab, 1)  

```

We performed a rowwise evaluation and are now able to answer the following question: 

* Approximately what proportion of all industrial workers are white? 
* The answer is: around 86%. 


## Chi-squared Test of Independence

Finally, let's test the hypothesis whether the variable `jobclass` is independent of the variable `race` at .05 significance level.


```{r chi-sq}

chisq.test(tab)  

```

As the p-value is smaller than the .05 significance level, we reject the null hypothesis that the jobclass is independent of the race of the workers. 