# Barplot {#barplot}


:::note
You should use this method if the data is:

- Categorical
:::

In this chapter you will learn how to do some simple data explorations for categorical variables using barplots. 

## One variable

```{r bar-1}

wage_df %>% 
  ggplot(aes(x = maritl)) +
  geom_bar()

```


## Two variables

### Stacked barplot

Absolute values:

```{r bar-stacked}

wage_df %>% 
  ggplot(aes(x = maritl, fill = education)) + 
  geom_bar()

```

Relative values: 

```{r bar-stacked-rel}

wage_df %>% 
  ggplot(aes(x = maritl, fill = education)) +
  geom_bar(position = "fill") +
  ggtitle("Marital Status", "Overview") +
  xlab(" Marital Status") +
  ylab("Number of People") +
  theme_classic() +
  scale_fill_brewer(palette = "Blues") +
  theme(legend.title = element_blank())

```


### Side-by-side barplot

Basic plot:

```{r bar-side}

wage_df %>% 
  ggplot(aes(x = maritl, fill = education)) +
  geom_bar( position = "dodge") 

```

Plot with some adjustments:


```{r bar-side-adj}

wage_df %>% 
  ggplot(aes(x = maritl, fill = education)) +
  geom_bar( position = "dodge") +
  ggtitle("Marital Status", "Overview") +
  xlab(" Marital Status") +
  ylab("Number of People") +
  theme_classic() +
  scale_fill_brewer(palette = "Blues") +
  theme(legend.title = element_blank())

```

### Faceted barplot

Basic plot:

```{r bar-facet}

wage_df %>% 
  ggplot(aes(x = maritl)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ education)

```


Plot with some adjustments to change our x labels using [justifications](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html)):

Horizontal and vertical justification have the same parameterisation, either a string (“top”, “middle”, “bottom”, “left”, “center”, “right”) or a number between 0 and 1:"

* top = 1, middle = 0.5, bottom = 0
* left = 0, center = 0.5, right = 1


```{r bar-favet-adj}

wage_df %>% 
  ggplot(aes(x = maritl)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ education) + 
  ggtitle("Marital Status", "Overview") +
  xlab(" Marital Status") +
  ylab("Number of People") +
  theme_classic() +
  theme(legend.title = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust=  1))

```
